Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-04 paper/paper.tex
pass

\section{Illustrative Example}
\label{sec:suri-metas-coverage}
\label{sec-:active-recon}

@@

\tname{} is a technique to synthesize rules for signature-based
\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
goal is to create rules that capture \emph{only} the
malicious traffic.

\vspace{1ex}
\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
benign and malicious network traffic and (2) rules from public
rulesets. \tname\ takes on input malicious and benign traffic and a
set of rules from public rulesets and produces on output a list of
\emph{plausible rules}---\ie{}, rules that isolate the malicious
traffic from the benign traffic---ranked in order of similarity with
existing rules. Figure~\ref{fig:overview} shows the workflow of
\tname{} as a pipeline of three components, which appear numbered in
the figure.  Public data sets of benign
traffic~\cite{tcpreplay,stratosphere-normal} and \nids\ rules
exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
only needs to provide the malicious traffic on input.

\begin{figure}[h!]
  \centering
  \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
  \caption{The \tname\ workflow.\Mar{revisar}}
  \label{fig:overview}
  \vspace{-2ex}
\end{figure}

\tname\ works as follows. First, it produces a rule that captures the
malicious traffic. In this step, \tname\ detects options directly from
the input messages (see Section~\ref{sec:rules-and-packets}). More
precisely, it produces a rule containing all possible options
satisfied by the traffic. Let us refer to that rule as the \emph{seed
  rule} and its corresponding set of options as \emph{seed
  options}. Note that, as only the malicious traffic is considered in
this step, the seed rule is likely overspecified, \ie{}, it may be
more restrictive than the golden rule. Consequently, the use of that
rule could result in false negatives. Second, \tname\ generates
alternative rules that 1) weaken the constraints expressed in the seed
rule while 2) still capturing the malicious traffic. Given that every
option declared in a rule needs to be satisfied for a rule to capture
an attack, any subset of the set of seed options, by definition,
captures the attack characterized by the malicious traffic. \tname{}
leverages that observation to search for rules containing a proper
subset of options from the seed options. However, note that that
approach alone could result in false positives as it imposes no limit
on to what extent a rule can be relaxed. \tname{} uses benign traffic
for that purpose. More precisely, it uses a gradient descent-like
search to find plausible rules derived from the seed rule, obtained in
the previous step. Finally, \tname{} uses several heuristic functions
derived from a data set of existing rules to rank the generated
plausible rules by their similarity with existing rules. Similarity is
used to approximate user's intent.

@@

The first step in the \tname's workflow is Reverse Engineering. The
goal of this step is to extract options from the malicious
traffic. Consequently, a rule that contains these options captures the
malicious traffic. \tname\ uses different strategies to extract
options based on the protocol detected in the input traffic: 1)~the
field strategy applies to any protocol and 2)~the payload strategy
applies only to http. We elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the network message field
data. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri and
Section~\ref{sec:active-recon} shows a concrete example where the
values for options are directly extracted from the fields of
messages. We obtained the option-field mapping manunually, from the
\suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload; not the fields. We found empirically that this occurs
frequently in http attacks, which are prevalent. For example,
\percHttp\ of the \numrulessuri{} rules from a public data
set~\cite{emerging-threats-open} we analyzed prescribe defenses
against http attacks and the majority of these rules use content
options. Figure~\ref{fig:rule-jboss} shows an example of a rule that
captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
payload.  Content options are very common in http
rules---\percRulesWithContent{} of the rules from the data set
mentioned above contain at least one content
option. Figure~\ref{fig:distribution-contents} shows the distribution
of number of content options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 in the
histogram for space. \MyComment{Rules without \CodeIn{content} options
  are an exception and half of the rules contain at least two
  \CodeIn{content} options.}  To sum up, handling content options for
http is very important.

The content strategy focuses on extracting content options from the
message payload. Inspecting existing rules, we found that these
options refer, in the payload, to sequences of characters separated by
typical natural language delimiters such as $\backslash$t,
$\backslash$n, $\backslash$r, and spaces. To find the values
associated with these options, \tname{} splits the payload of the
malicious message in tokens using these delimiters.  Considering the
example from Section~\ref{sec:content-example}, \tname{} produced a
total of 23 tokens, considered content option candidates.

It is worth noting that we found empirically that deciding which
reverse engineering strategy to use based on the protocol of the input
(malicious) messages was effective. We used the payload strategy for
http and the field strategy for the other protocols. For flooding
attacks, \tname\ runs the same approach for each message and
identifies common options across all messages. For that case,
\tname\ reports, in addition, a \CodeIn{threshold} option, as
discussed on Section~\ref{sec:rules-and-packets}.  At the end of the
Reverse Engineering step, \tname\ produces a seed rule that captures
the malicious traffic.

>>>>

\section{Illustrative Example}
\label{sec:suri-metas-coverage}
\label{sec:active-recon}

@@

\tname{} is a technique to synthesize rules for rule-based
\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
goal is to automatically create rules that capture \emph{only} the
malicious traffic.

\vspace{1ex}
\noindent\textbf{Overview.}~\tname\ takes on input one or more
examples of attacks and produces on output a ranked list of
\emph{plausible rules}---\ie{}, rules that isolate the malicious
traffic produced by the attacks from the benign
traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
pipeline of three components, appearing numbered in the figure.
Public data sets of benign
traffic~\cite{tcpreplay,stratosphere-normal} and rules
exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
only needs to provide examples of attacks on input.

%% To create rules, \tname\ leverages (1)
%% benign and malicious network traffic and (2) rules from public
%% rulesets.

\begin{figure}[t!]
  \centering
  \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
  \vspace{-4ex}  
  \caption{The \tname\ workflow.\Mar{revisar}}
  \label{fig:overview}
  \vspace{-3ex}
\end{figure}

\tname\ works as follows. First, it creates a rule that captures the
malicious traffic produced by the examples provided on input. In this
step, \tname\ detects options directly from the input messages (see
Section~\ref{sec:rules-and-packets}). More precisely, it creates a
rule containing options that are guaranteed to be satisfied by the
malicious traffic. Let us refer to that rule as the \emph{seed rule}
and its corresponding set of options as \emph{seed options}. Note
that, as only the malicious traffic is considered in this step, the
seed rule is likely overspecified, \ie{}, it may be more restrictive
than the golden rule. Consequently, the use of that rule could result
in false negatives. Second, \tname\ generates alternative rules that
1) weaken the constraints expressed in the seed rule while 2) still
capturing the malicious traffic. Given that every option declared in a
rule needs to be satisfied for a rule to capture an attack, any subset
of the set of seed options, by definition, captures the attack
characterized by the malicious traffic. \tname{} leverages that
observation to search for plausible rules containing a proper subset
of options from the seed options. However, note that that approach
alone could result in false positives as it imposes no limit on to
what extent a the constraints induced from the rule can be
weakened. \tname{} uses benign traffic for that purpose. More
precisely, it searches for plausible rules derived from the seed rule
obtained in the previous step. Finally, \tname{} uses heuristic
functions obtained from a data set of existing rules to rank the
generated plausible rules by their similarity to existing
rules. Similarity is used to approximate user's intent. The following
sections detail each component of \tname.

@@

The goal of the Reverse Engineering step is to extract options from
the malicious traffic produced by the input example attacks. A rule
that contains these options captures the malicious traffic by
construction. \tname\ uses two strategies to extract options: 1)~the
field strategy applies to any protocol and 2)~the payload strategy
applies only to http.  The decision of which strategy to use is based
on the protocol associated with input traffic--the technique uses the
payload strategy if http is detected; otherwise, it uses the field
strategy. We elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of
messages. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri\ and
Section~\ref{sec:active-recon} shows a concrete example illustrating a
case of this strategy. We obtained the option-field mapping
manunually, from the \suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload; not the fields. By analyzing our data set of rules, we found
that this occurs frequently in http attacks, which are prevalent. For
example, \percHttp\ of the \numrulessuri{} rules from the ruleset we
analyzed~\cite{emerging-threats-open} prescribe defenses against http
attacks. In addition, \percRulesWithContent{} of these rules contain
at least one content option. Figure~\ref{fig:rule-jboss} shows an
example of a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.
Figure~\ref{fig:distribution-contents} shows the distribution of
number of content options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 in the
histogram for space. \MyComment{Rules without \CodeIn{content} options
  are an exception and half of the rules contain at least two
  \CodeIn{content} options.}  To sum up, handling content options for
http is very important.

The content strategy focuses on extracting content options from the
message payload. By inspecting existing rules, we found that these
options refer to sequences of characters separated by typical natural
language delimiters such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. To find the values associated with these
options, \tname{} splits the payload in tokens using these delimiters.
Considering the example from Section~\ref{sec:content-example},
\tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
  Se sim, como isto cai para 5 (pela Tabela 3)}, considered content
option candidates.

It is worth noting that we found that deciding which reverse
engineering strategy to use based on the protocol of the input
(malicious) messages was effective. We used the payload strategy for
http and the field strategy for the other protocols. For flooding
attacks, \tname\ runs the same approach for each message and
identifies common options across all messages. For that case,
\tname\ reports, in addition, a \CodeIn{threshold} option, as
discussed on Section~\ref{sec:rules-and-packets}.  At the end of the
Reverse Engineering step, \tname\ produces a seed rule that captures
the malicious traffic.
